### 2016-10-07 robust-software.tex
Hopefully finalizing changes

\begin{flushleft}
{\Large
\textbf\newline{Ten Simple Rules for Making Research Software More Robust}
}
\newline
\\
{Morgan~Taschuk}\textsuperscript{1,\ddag *},
{Greg~Wilson}\textsuperscript{2}
\\
\textbf{1} Ontario Institute for Cancer Research / morgan.taschuk@oicr.on.ca
\\
\textbf{2} Software Carpentry Foundation / gvwilson@software-carpentry.org
\\
\bigskip
{\ddag} These authors contributed equally to this work.
\\
* Corresponding author.
\end{flushleft}

@@

Typically, scientific software is developed and used by a single person,
usually a graduate student or postdoc, and produces intended results in their
hands~\cite{prins2015}. 
But what happens
when someone else wants to use their software? Everyone with
a few years of experience feels a bit nervous when told to use
a graduated student's code to analyze their data.
Often, that software will be undocumented and work in unexpected
ways (if it works at all).  It will often rely on nonexistent paths or resources,
be tuned for a single dataset,
or simply be an older version than was used in published papers.
The new user is then faced with two unpalatable options:
hack the existing code to make it work, or start over. Being unable to
replicate results is so common that one publication refers to it as "a rite of
passage"~\cite{baker2016}.

The root cause of this problem is that most research software
isn't \emph{robust}. The difference between running and being
robust is the difference between ``works for me on my machine'' and
``works for other people on a cluster I've never used''. With a lack of
robustness also comes a lack of reproducibility and many duplicated efforts,
slowing the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics is full of such efforts, catalogued by efforts such as
the ELIXIR tools and data registry~\cite{ison2016}
and the Bioinformatics Links Directory~\cite{brazas2012}. In the Bioinformatics
Links Directory as of 2016, there are 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.

This problem is not unique to bioinformatics, or even to computing~\cite{baker2016}. 
Best practices in software engineering specifically aim to increase software
robustness. However, most bioinformaticians come from related disciplines and
are trained on-the-job or otherwise informally~\cite{prins2015,atwood2015}. Even
the existing programs and initiatives rarely have the time to cover software engineering
in-depth, especially since the field is so broad and rapidly
developing~\cite{atwood2015,lawlor2015}. Regardless of whether or not the skills
and knowledge are taught, developing robust software is not directly rewarded
in science and funding is difficult to come by~\cite{prins2015}. Some proposed
solutions to this problem, in addition to restructuring educational programs,
include hiring dedicated software engineers~\cite{lawlor2015,sanders2008},
partnering with private sector or grassroots organizations~\cite{prins2015,ison2016},
or using specific technical tools like containerization or cloud
computing~\cite{afgan2016,howe2012}. Each of these requires time and, in some
cases, institutional change.

@@

That said, not \emph{every} coding effort requires such rigor:
as the saying goes, not everything worth doing is worth doing right (or right away).
Code that is used once to answer a specific question related to a specific dataset
doesn't require comprehensive documentation or flexible configuration,
and the only sensible way to test it may well be to run it on the dataset in
question. Exploratory analysis is an iterative process that is developed
quick and revised often~\cite{lawlor2015,sanders2008}. 
However, if a script is dusted off and run three or four
times for slightly different purposes,
is crucial to a publication or a lab,
or being passed on to someone else,
it may be time to make your software more robust.
In particular, robust software:

@@

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it.
A working example using test data (per
\#9) is always appreciated.

@@

\textbf{List required dependencies:} Often, software depends on very
specific versions of libraries, modules, or operating systems. This is
entirely reasonable as long as it is properly documented. Often,
multiple libraries exist with the same or very similar names, so
either provide the commands necessary to download the dependencies or
link to the software homepage. Include the version number for each
dependency.

If your dependency is not publicly available,
you have several options to ensure your software remains useable.
You can add plain-text or code directly to
your repository with appropriate attribution. If the dependency is a binary, use a
\withurl{binary repository manager}{https://en.wikipedia.org/wiki/Binary\_repository\_manager}
such as Artifactory or ProGet. These managers keep versioned copies of
software at constant URLs so they can be downloaded as long as the
manager continues to run. If the dependency must remain closed, place it at an
internal location on shared disk, remove all write permissions, and
link to it from your README, although this method is 
discouraged because of the potential for sharing and risks accidental removal.

@@

\textbf{Attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how you want others to use and
credit your software. By putting them in the README, it is readable \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of ``expert'' users
that can be contacted if new users have problems with the software.

The README file for \withurl{khmer}{https://github.com/dib-lab/khmer/blob/master/README.rst}
is a good model:
it explains the software's purpose,
tells readers where to get help,
and includes links to a CITATION file (explaining how to cite the project)
and the license.

@@

An example of good usage is GNU's \texttt{mkdir} command, which makes
new directories:

@@

There is no standard format for usage statements, but good ones share
several features:

@@

\section{Give the software a meaningful version number.}

Most software has a version number composed of a decimal number that
increments as new versions are released. There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
Every time you release your software, i.e., distribute it to
someone other than yourself and/or the development team, you should
increment your version number.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g., 0.2.6. The major and
minor numbers versions are almost always provided. Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software, like adding new
features. Following the minor version number can be an arbitrary number
of project-specific identifiers, including patches, builds and qualifiers.
Common qualifiers include \texttt{alpha}, \texttt{beta}, and \texttt{SNAPSHOT},
for applications that are
not yet stable or released, and \texttt{-RC} for release candidates prior
to official release.

The version of your software should be easily available, both when
supplying \texttt{-\/-version} or \texttt{-v} on the command line as
well as in the results. The software version should be printed to the
same location as all of the other parameters (see Rule 4).  It should
also be included in all of the program's output, particularly debugging
traces: if someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

\section{Make older versions available.}

Software evolves, but not all software evolves at the same pace.
While more consistent command-line arguments
or more flexible handling of output directories
may make a program better in general,
it can simultaneously make work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore ensure that every version they have released
continues to be available.

A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control, to official releases alongside
code on Sourceforge, Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

The most common way to do this is to include the version number in the
name of the released package, and to link to those old versions from
the project's home page or downloads page.
For example,
as of the time of writing,
the downloads page for the Python programming language includes links to
Versions 3.5.2, 3.5.1, 3.5.0, 3.4.5, 3.4.4, 2.7.12, and 2.7.11,
along with a link to an archive page of even older versions.
Such archives are most useful if these packages document their dependencies,
which we discuss in the next section.

@@

In the spirit of code reuse and interoperability, developers often want
to reuse software written by others. 
With a few lines, a call
is made out to the other program and the results are incorporated into the
primary script. Using popular projects reduces the amount of code that
needs to be maintained and leverages the work done to check that software.

Unfortunately, reusing software introduces dependencies, which can
bring their own special pain.  The interface between two software
packages can also be a source of considerable frustration: all too
often, support requests descend into debugging errors produced by the
other project.

In addition, every package someone has to install
before being able to use yours is a
possible (some would say ``likely'') source of frustration for some
potential user. On the other hand, research software developers should
re-use existing software wherever possible.

@@

The way the second program is invoked can throw up errors.
The program may not be in the user's path, or it may be an
older or newer version and produce results different than expected.
Windows users will be frustrated if you invoke \texttt{bash} or \texttt{sh} explictly or
use shell-specific conventions like `*' expansion.
Even Linux-standard functions available vary slightly between
installs. For example, GNU \texttt{sort} is available on almost every
*nix distribution, but sorts differently depending on locale.

Despite these caveats,
we fully support the reuse of software between projects provided
three additional guidelines are adhered to. First, 
\textbf{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration.

Second, \textbf{ensure the appropriate software and version is available}.
Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using a
dependency management system. Regardless, check whether the program is
executable and what version is running. Be sure to remind users in the
documentation what the compatible versions are.

@@

Another reason for this rule is that users may want to try out a new
package before installing it system-wide on a cluster. Requiring root
privileges will frustrate such efforts, and thereby reduce uptake of the
package. Requiring that software be installed under its own user account
(e.g., that \texttt{packagename} be made a user, and all of the
package's software be installed in that ``user's'' space) is similarly
limiting, and makes side-by-side installation of multiple versions of
the package more difficult.

@@

Note:
testing software installation has traditionally been regarded as difficult,
since it necessarily alters the machine on which the test is conducted.
Lightweight virtualization containers like Docker make this much easier as well.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but a short list includes
input and reference files and directories,
output files and directories,
filtering,
tuning,
random number generation seeds,
and
any alternatives that you've built-in,
e.g.,
compress results, use a different calculation, verbose output, etc.

Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal seed for your
original dataset, but that doesn't mean that is the best seed for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
facilitating exploratory analysis and parameter sweeping. Keep in mind
that if a parameter can be adjusted, users will want to be able to turn
off the feature entirely to have a baseline comparison.

When the software starts, it should echo all parameters and software
versions to standard out or a log file alongside the results. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

You can set reasonable default values
as long as any command line arguments
override those values. Configuration files should be used in preference
to hardcoding the defaults directly. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.

We caution against overusing configuration files. When a user needs to
locate, open, change and save a file in order to change a parameter, the
import of the change seems larger and discourages experimentation. 
Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institution. 
Since
all parameters should be echoed in the results, config files should be
cleaned up after execution completes or they occupy valuable disk space.
Use a default configuration file and specify all other parameter values on the
command line.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is \emph{not} to check that the software
is working correctly (although that is extremely helpful), but rather to
ensure that it will work at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, this test script must be easy to find and run. A
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them into the shell.

@@

Finally,
studies have found that the ease with which people can start making
contributions is a strong predictor of whether they will or not~\cite{steinmacher2015}.
By making it simpler for outsiders to contribute,
a test suite of any kind also makes it more likely that they will.

@@

Many applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility impossible.
Debugging is more difficult. If even the small test set (\#9) produces
different results, new users may not be able to tell whether the software is
working properly. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

Given the size of biological data, it is unreasonable to suggest that
random algorithms be removed. However, most programs use a pseudo-random
number generator, which uses a starting seed and an equation to
approximate random numbers. Setting the seed to a consistent value
removes randomness between runs. Allow the user to optionally provide
the seed as an input parameter, thus rendering the program deterministic
for those cases where it matters. If the seed is set internally (e.g.,
using clock time), echo it to the output for re-use later.

@@

\item[\textbf{Documentation}] We think it is very important for
  developers to document their work, but our experience is that people
  are unlikely do it during normal development, and we don't want to
  recommend practices that people won't actually adopt.

\item[\textbf{Build Systems}] If the software depends on a build tool,
  either because it is written in a compiled language or to manage
  workflow, it should use a standard tool such as Make rather than
  home-brewed scripts.

\end{description}

>>>>

\begin{flushleft}
{\Large
\textbf\newline{Ten Simple Rules for Making Research Software More Robust}
}
\newline
\\
{Morgan~Taschuk}\textsuperscript{1,\ddag *},
{Greg~Wilson}\textsuperscript{2,\ddag}
\\
\textbf{1} Ontario Institute for Cancer Research / morgan.taschuk@oicr.on.ca
\\
\textbf{2} Software Carpentry Foundation / gvwilson@software-carpentry.org
\\
\bigskip
{\ddag} These authors contributed equally to this work.
\\
* Corresponding author.
\end{flushleft}

@@

Typically, scientific software is developed and used by a single person,
usually a graduate student or postdoc, and produces the intended results in their
hands~\cite{prins2015}. 
But what happens
when someone else wants to use their software? Everyone with
a few years of experience feels a bit nervous when told to use
another person's code to analyze their data.
Often, that software will be undocumented and work in unexpected
ways (if it works at all).  It will often rely on nonexistent paths or resources,
be tuned for a single dataset,
or simply be an older version than was used in published papers.
The new user is then faced with two unpalatable options:
hack the existing code to make it work, or start over. Being unable to
replicate results is so common that one publication refers to it as "a rite of
passage"~\cite{baker2016}.

The root cause of this problem is that most research software
isn't \emph{robust}. The difference between running and being
robust is the difference between ``works for me on my machine'' and
``works for other people on a cluster I've never used''. With a lack of
robustness also comes a lack of reproducibility and many duplicated efforts,
slowing the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics is full of such efforts, catalogued by repositories such as
the ELIXIR tools and data registry~\cite{ison2016}
and the Bioinformatics Links Directory~\cite{brazas2012}. In the Bioinformatics
Links Directory as of 2016, there are 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

This problem is not unique to bioinformatics, or even to computing~\cite{baker2016}. 
Best practices in software engineering specifically aim to increase software
robustness. However, most bioinformaticians come from related disciplines and
are trained on-the-job or otherwise informally~\cite{prins2015,atwood2015}. Even
the existing training programs and initiatives rarely have the time to cover software engineering
in-depth, especially since the field is so broad and rapidly
developing~\cite{atwood2015,lawlor2015}. Regardless of whether or not the skills
and knowledge are taught, developing robust software is not directly rewarded
in science and funding is difficult to come by~\cite{prins2015}. Some proposed
solutions to this problem, in addition to restructuring educational programs,
include hiring dedicated software engineers~\cite{lawlor2015,sanders2008},
partnering with private sector or grassroots organizations~\cite{prins2015,ison2016},
or using specific technical tools like containerization or cloud
computing~\cite{afgan2016,howe2012}. Each of these requires time and, in some
cases, institutional change.

@@

That said, not \emph{every} coding effort needs to be engineered to last.
Code that is used once to answer a specific question related to a specific dataset
doesn't require comprehensive documentation or flexible configuration,
and the only sensible way to test it may well be to run it on the dataset in
question. Exploratory analysis is an iterative process that is developed
quick and revised often~\cite{lawlor2015,sanders2008}. 
However, if a script is dusted off and run three or four
times for slightly different purposes,
is crucial to a publication or a lab,
or being passed on to someone else,
it may be time to make your software more robust.

In particular, robust software:

@@

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards}. We summarize the crucial points here.

@@

\textbf{List required dependencies:} Often, software depends on
specific versions of libraries, modules, or operating systems. 
This is
entirely reasonable as long as it is properly documented. Include the name and
version number for each dependency.  Several methods exist for acquiring
required packages: either provide the commands necessary to download the
dependencies; link to the software homepage; or (our preference) use a
dependency manager. We address dependencies in more detail in rule
\textcolor{red}{5}.

@@

\textbf{Attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how you want others to use and
credit your software. READMEs are usualy available \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of ``expert'' users
that can be contacted if new users have problems with the software.

The README file for \withurl{khmer}{https://github.com/dib-lab/khmer/blob/master/README.rst}
is a good model:
it explains the software's purpose,
tells readers where to get help,
and includes links to a CITATION file (explaining how to cite the project)
and the license file.

@@

Almost all command-line applications use a combination of
POSIX~\cite{posix2016} and GNU~\cite{gnustandards} standards for usage. 
An example of good usage is GNU's \texttt{mkdir} command, which makes
new directories:

@@

@@

\section{Version your releases}

Software evolves over time, with developers adding or removing features as need
dictates. Making official releases stamps a particular set of features with a
project-specific identifier so that version can be retrieved for later use. For
example, if a paper is published, the software should be released at the same
time so that the results can be reproduced. 
Most software has a version number composed of a decimal number that
increments as new versions are released.

There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
\textbf{Increment your version number every time you release your software to
other people}.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g., 0.2.6.
Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software, like adding new
features. Following the minor version number can be an arbitrary number
of project-specific identifiers, including patches, builds and qualifiers.
Common qualifiers include \texttt{alpha}, \texttt{beta}, and \texttt{SNAPSHOT},
for applications that are
not yet stable or released, and \texttt{-RC} for release candidates prior
to an official release.

The version of your software should be easily available.
Supplying \texttt{-\/-version} or \texttt{-v} on the command line should print
the software name and version number, and it should
also be included in all of the program's output, particularly debugging
traces: if someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

While new releases may make a program better in general,
they can simultaneously create work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore ensure that every version they have released
continues to be available. A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control, to official releases alongside
code on Sourceforge, Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

@@

In the spirit of code reuse and interoperability, developers often want
to reuse software written by others. 
With a few lines, a call
is made out to the other program and the results are incorporated into the
primary script. Using popular projects reduces the amount of code that
needs to be maintained and leverages the work done by the other software.

Unfortunately, reusing software introduces dependencies, which can
bring their own special pain. The interface between two software
packages can also be a source of considerable frustration: all too
often, support requests descend into debugging errors produced by the
other project. In addition, every package someone has to install
before being able to use yours is a
possible (some would say ``likely'') source of frustration for some
potential user. The way the second program is invoked can throw up errors.
The program may not be in the user's path, or it may be an
older or newer version and produce results different than expected.
Windows users will be frustrated if you invoke \texttt{bash} or \texttt{sh} explictly or
use shell-specific conventions like `*' expansion.
Even Linux-standard functions available vary slightly between
installs. For example, GNU \texttt{sort} is available on almost every
*nix distribution, but sorts differently depending on locale.

Despite these problems, software developers in research should
re-use existing software wherever possible and reasonable to avoid
``reinventing the wheel''. We fully support the reuse of software between
projects provided three additional guidelines are adhered to.

First, 
\textbf{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration.

Second, \textbf{ensure the appropriate software and version is available}.
Early in execution, check whether the dependency is executable and whether the
version is compatible.
Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using a
dependency management system (see Rule \textcolor{red}{RULENO}).

\section{Use a build system and dependency manager.}

If the software depends on a build tool, either because it is written in a
compiled language or to manage a workflow, it should use a standard tool such
as Make rather than home-brewed scripts.

@@

@@

Another reason for this rule is that users may want to try out a new
package before installing it system-wide on a cluster. Requiring root
privileges will frustrate such efforts, and thereby reduce uptake of the
package. Requiring that software be installed under its own user account
(e.g., that \texttt{packagename} be made a user, and all of the
package's software be installed in that ``user's'' space (see: Apache Tomcat))
is similarly
limiting, and makes side-by-side installation of multiple versions of
the package more difficult.

@@

Testing the ability to install software has traditionally been regarded as difficult,
since it necessarily alters the machine on which the test is conducted.
Lightweight virtualization containers like Docker make this much easier as
well, or simply asking another person to try and build your software.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application.
The full list of useful parameters is software-specific and so cannot be
detailed here, but a short list includes providing
input and reference files and directories,
output files and directories,
filtering,
tuning,
random number generation seeds,
and
any alternatives that you've built-in,
e.g.,
compress results, use a different calculation, verbose output, etc.

\textbf{Allow the user to configure useful parameters on the command line.}
Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal seed for your
original dataset, but that doesn't mean that is the best seed for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
facilitating exploratory analysis and parameter sweeping.

When the software starts, it should \textbf{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

You can set reasonable default values
as long as any command line arguments
override those values. Configuration files should be used in preference
to hardcoding the defaults directly. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.
Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institution. 
Since
all parameters should be echoed in the results, config files should be
cleaned up after execution completes or they occupy valuable disk space.
Use a default configuration file and specify all other parameters on the
command line.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is not only to check that the software
is working correctly (although that is extremely helpful), but rather to
ensure that it will work at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, this test script must be easy to find and run. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them into the shell.

@@

Finally,
studies have found that the ease with which people can start making
contributions is a strong predictor of whether they will or not~\cite{steinmacher2015}.
By making it simpler for outsiders to contribute,
a test suite of any kind also makes it more likely that they will, and software
with collaborators stands a better chance of surviving in the busy field of
scientific software.

@@

Many applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility and therefore
debugging more difficult. If even the small test set (\#9) produces
different results, new users may not be able to tell whether the software is
working properly. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

Given the size of biological data, it is unreasonable to suggest that
random algorithms be removed. However, most programs use a pseudo-random
number generator, which uses a starting seed and an equation to
approximate random numbers. Setting the seed to a consistent value
can remove randomness between runs. Allow the user to optionally provide
the seed as an input parameter, thus rendering the program deterministic
for those cases where it matters. If the seed is set internally (e.g.,
using clock time), echo it to the output for re-use later. 
If setting the seed is not possible, make sure the acceptable tolerance is
known and detailed in documentation and in the tests.

@@

\item[\textbf{Documentation}] We think it is very important for
  developers to document their work, but our experience is that people
  are unlikely do it during normal development, and we don't want to
  recommend practices that people won't actually adopt. However, it is worth
  noting that software that is widely used and contributed to has and enforces
  the need for good documentation~\cite{gentleman2004}.

\item[\textbf{Build Systems}] 
\end{description}
